Constraining Black Hole Growth with High-z Quasar Observations

### 2022-07-31 ms.tex
version 0731; intro revised by k; need add a SMBH track crossing Wang2021, QLF result and detection limit description

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of black hole mass function (BHMF) and the quasar luminosity function (QLF)
plays an important role in the establishment of BH-galaxy coevolution at high redshifts ($z\gtrsim$ 6).
However, understanding the accretion and radiation properties of growing BHs in the high-$z$ environments 
still remains challenging.
In this work, based on a semi-analytic BH seed formation model, 
we relate BH seeds with the growth tracks of their parent halos. 
We then construct a model of BH mass and luminosity evolution 
characterizing the episodical BH accretion and Eddington ratio distribution function (ERDF).
The model parameters are fitted to reproduce the observed BHMF and QLF of the quasar population at $z\sim$ 6.
Our best-fit models suggest that BHs grow with episodical bursts and quescent phases, with time duration 
typically in units of 20 $\sim$ 30 Myr. 
The duty cycle of active accretion is $f_{\rm duty}\sim 0.15$ for massive BHs with $M\gtrsim 10^8~\Msun$, 
boosted from that in the whole population by nearly a factor of two. 
We find that the intrinsic ERDF of the whole quasar population is well presented by a Schechter shape, 
indicating a large number of low Eddington ratio objects missed by the current flux-limited quasar selections.
We give prediction of the early evolution of BHMF and QLF,
as well as the growth patterns of individual BHs, 
which will likely be unveiled by the James Webb Space Telescope (JWST) and future deep and wide-field surveys.
%  with, e.g. 
% the , Roman Space Telescope (RST), and Euclid.
\end{abstract}

@@

\vspace{5mm}
\section{Introduction} \label{sec:intro}
The growth of massive black holes (BHs) in active galactic nuclei (AGN) is elegantly related to their luminous accretion phases
across the cosmic time \citep{1982MNRAS.200..115S}.
%,1992MNRAS.259..725S}.
Comparing the mass density of massive BHs in the local universe and the mass accreted onto BH inferred from
the integration of the quasar luminosity function based on multi-wavelength observations,
the cosmic BH accretion history and their radiative efficiency are nicely constrained 
\citep[e.g.][]{1971ApJ...170..223C,1992MNRAS.259..725S,2002MNRAS.335..965Y,2004MNRAS.351..169M,2008MNRAS.388.1011M,2004MNRAS.354.1020S,
2009ApJ...690...20S,2014MNRAS.439.2736D,2014ApJ...786..104U,2017A&A...600A..64T}.

Extrapolation of this argument toward higher redshifts beyond $z\sim$ 6 is limited because of the current capability for
high-$z$ quasar observations \citep{2013ApJ...768..105M,2016ApJ...829...33Y,2019ApJ...884...30W,2019BAAS...51c.121F}.
% \red{KI: you could find better reference in this Decadal Survey?}.
Nevertheless, currently the $z\sim 6$ quasar population is well constrained by multiband surveys
\citep{2008AJ....135.1057J,2010AJ....139..906W,2016arXiv161205560C,2016ApJ...833..222J,2018PASJ...70S..35M,2019ApJ...883..183M,2019AJ....157..168D}
along with the quasar luminosity function \citep[e.g.,][]{2018ApJ...869..150M}.
%and the virial BH mass estimates
%from the \ion{Mg}{2} 2798 $\rm{\AA}$ line width 
%where 
Deep spectroscopic observations of high-$z$ quasars enable us to measure the virial BH mass 
with the \ion{Mg}{2} single-epoch method and bring insights of the mass distribution
\citep{2007AJ....134.1150J,2007ApJ...669...32K,2010AJ....140..546W,2018Natur.553..473B,2019ApJ...880...77O,2019ApJ...873...35S}. 
Moreover, in the era of the James Webb Space Telescope (JWST) and forthcoming facilities,
e.g., the Roman Space Telescope (RST) and Euclid,
infrared imaging and spectroscopic observations will unveil a wealth of information on the
high-$z$ quasar properties and their environments 
\citep{2019BAAS...51c..45R, 2019arXiv190205569A, 2011arXiv1110.3193L}. 
Deep surveys of high-$z$ quasars and their host galaxies will shed light on the early BH evolution, 
also help answering the intriguing questions e.g., 
the existence of supermassive BHs in the early universe within one billion year 
\citep{2012Sci...337..544V,2013ASSL..396..293H,2020ARA&A..58...27I}, 
and the establishment of BH-galaxy coevolution 
\citep[e.g.,][]{2013ApJ...773...44W,2017ApJ...851L...8V,2021ApJ...914...36I,2022ApJ...927..237I,2022MNRAS.511.3751H}.
% \red{cite Wang+2013; Venemans+2017, Izumi et al. 2021, Inayoshi+2022 simulation, Habouzit+2022}.

Formation of seed BHs is the starting point to characterize the early assembly of massive BH populations.
Their formation pathway and expected mass range have been extensively investigated by numerical simulations and
semi-analytical studies
%From Pop III star remnants ($10-10^3 ~\Msun$) to direct collapse supermassive stars (10$^4-10^5 ~\Msun$),
%the properties of high-$z$ BH seeds are extensively modeled by simulations and semi-analytical works
%\citep[e.g.,][Toyouchi 2022]{2001ApJ...546..635O,2002Sci...295...93A,2006ApJ...652....6Y,2011MNRAS.416.2748I,
%2012ApJ...756...93H,2013ApJ...778..178H,2014MNRAS.439.3798F,
%2014MNRAS.445L.109I,2014MNRAS.445..544S,2014MNRAS.445..107V,2020MNRAS.499.5960S}.
\citep{2006MNRAS.370..289B,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T}.
% \red{KI: I think here we should cite more general citations: Begelman+2006; Tanaka \& Haiman 2009; Hirano+14, 15,
% Toyouchi+2022, one Natarajan+Volonteri's paper discussing the mass function, Valiante+2018, Sassano+21}
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
%factors which modulate the Pop III star formation have been assessed,
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
%irradiation in Lyman-Werner band which photo-dissociates the major coolant $\hh$
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
% \red{KI: add Chon+16},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
% \red{KI: add Inayoshi, Li \& Haiman 18, Schauer +17},
and
(iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
% \red{KI: add Latif's Nature paper, Mayer+10 or 15}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},% \red{KI: add Becerra+15, Latif+16}, 
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
% \red{cite Hosokawa+13, Woods+19?, Schleicher+13, Toyouchi+22}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
% \red{cite Dijkstra+08, Ahn+09, Inayoshi \& Tanaka 15}, 
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the heavy-seed formation
can be accelerated in the rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and the frequency of halo mergers.
%These factors prevail in some environments at high redshifts and facilitate massive Pop III and even supermassive star formation,
%which contribute to the heavy seed scenario of supermassive BH formation at $z\gtrsim$ 6
%\citep[see a review by][]{2020ARA&A..58...27I}.

\red{Introduction of previous studies on QLFs for high-$z$ populations. mention Yung's paper, Ricarte \& Natarajan paper, recent Finkelstein's paper?
Also, I remember one paper of our former postdoc, Kim.
Also, recent Raffaella's student's paper and "highly-empirical paper" you introduced.
Please add more (I might miss some relevant papers.)}
\blue{
% Haiman \& Loeb 1997.
% High-z
}

To construct the picture of BH evolution, the initial mass function (IMF) of BH seeds is required.
Studies have shown that Pop III stars (or the seed BH population) formed in mini-halos is top-heavy, while capped with $\sim 10^3~\Msun$ 
\citep{2014ApJ...781...60H,2015MNRAS.448..568H}.
Focusing on the progenitor halos of supermassive BHs at $z\sim$ 6, 
our recent semi-analytical work produces an IMF of BH seeds ranging from $10^2-10^5~\Msun$ \citep{2021ApJ...917...60L}. 
The IMF is consistent with the three-dimensional radiation hydrodynamical simulations 
constituting the stellar evolution with large scale mass inflow under realistic halo circumstances (Toyouchi et al. 2022).
In this work, we apply the seeding model to examine the BH growth in common regions in the early universe, 
constrained by the observable QLF and BHMF down to $z\sim$ 6, 
% our aim is to understand the growth patterns of early BHs.

>>>>

%% Mark off the abstract in the ``abstract'' environment. 
\begin{abstract}
The early evolution of the quasar luminosity and black hole mass function (MF) encodes key information
on the physics determining the radiative and accretion processes of supermassive black holes (SMBHs) in high-redshift quasars.
Although the shape of the luminosity function (LF) has been constrained by recent wide and deep surveys,
it still remains challenging to develop a theoretical model that explains the redshift evolution and BH growth self-consistently.
In this study, based on a semi-analytical model for the formation and growth of seed BHs
in quasar progenitor halos,
we construct the LF and MF of the early BH population that experiences multiple accretion bursts,
in each of which the Eddington ratio is assigned following a Schechter function.
Our best-fit model to reproduce the observed LF and MF at $z\simeq 6$
suggests that several episodes of modest super-Eddington accretion occur
and each of them lasts $\tau \simeq 20-30$ Myr.
The average duty cycle of BHs that reach $\gtrsim 10^8~\Msun$ by $z\simeq 6$ is as high as $\simeq 15~\%$,
which is nearly twice that of the entire population.
We also find that the observed Eddington-ratio distribution function is skewed to a log-normal shape from the intrinsic Shechter function
owing to detection limits of quasar observations.
The predicted redshift evolution of both LF and MF of earliest BHs suggests a rapid decay of their number and mass
toward higher redshifts of $z>6$, compared to galaxy stellar mass.
This will be unveiled by future deep and wide surveys with the James Webb Space Telescope, Roman Space Telescope, and Euclid.
% the , Roman Space Telescope (RST), and Euclid.
\end{abstract}

@@

\vspace{5mm}
\section{Introduction} \label{sec:intro}
The growth of massive black holes (BHs) in active galactic nuclei (AGN) is elegantly related to their luminous accretion phases
across the cosmic time \citep{1982MNRAS.200..115S}.
%,1992MNRAS.259..725S}.
Comparing the mass density of massive BHs in the local universe and the mass accreted onto BH inferred from
the integration of the quasar luminosity function based on multi-wavelength observations,
the cosmic BH accretion history and their radiative efficiency are nicely constrained 
\citep[e.g.][]{1971ApJ...170..223C,1992MNRAS.259..725S,2002MNRAS.335..965Y,2004MNRAS.351..169M,2008MNRAS.388.1011M,2004MNRAS.354.1020S,
2009ApJ...690...20S,2014MNRAS.439.2736D,2014ApJ...786..104U,2017A&A...600A..64T,2022arXiv220607357S}.

Extrapolation of this argument toward higher redshifts beyond $z\sim$ 6 is limited because of the current capability for
high-$z$ quasar observations \citep{2013ApJ...768..105M,2016ApJ...829...33Y,2019ApJ...884...30W,2019BAAS...51c.121F}.
% \red{KI: you could find better reference in this Decadal Survey?}.
Nevertheless, currently the $z\sim 6$ quasar population is well constrained by multiband surveys
\citep{2008AJ....135.1057J,2010AJ....139..906W,2016arXiv161205560C,2016ApJ...833..222J,2018PASJ...70S..35M,2019ApJ...883..183M,2019AJ....157..168D}
along with the quasar luminosity function \citep[e.g.,][]{2018ApJ...869..150M}.
%and the virial BH mass estimates
%from the \ion{Mg}{2} 2798 $\rm{\AA}$ line width 
%where 
Deep spectroscopic observations of high-$z$ quasars enable us to measure the virial BH mass 
with the \ion{Mg}{2} single-epoch method and bring insights of the mass distribution
\citep{2007AJ....134.1150J,2007ApJ...669...32K,2010AJ....140..546W,2018Natur.553..473B,2019ApJ...880...77O,2019ApJ...873...35S}. 
Moreover, in the era of the James Webb Space Telescope (JWST) and forthcoming facilities,
e.g., the Roman Space Telescope (RST) and Euclid,
infrared imaging and spectroscopic observations will unveil a wealth of information on the
high-$z$ quasar properties and their environments 
\citep{2019BAAS...51c..45R, 2019arXiv190205569A, 2011arXiv1110.3193L}. 
Deep surveys of high-$z$ quasars and their host galaxies will shed light on the early BH evolution, 
also help answering the intriguing questions e.g., 
the existence of supermassive BHs in the early universe within one billion year 
\citep{2012Sci...337..544V,2013ASSL..396..293H,2020ARA&A..58...27I}, 
and the establishment of BH-galaxy coevolution 
\citep[e.g.,][]{2013ApJ...773...44W,2017ApJ...851L...8V,2021ApJ...914...36I,2022ApJ...927..237I,2022MNRAS.511.3751H}.

The shapes of the QLF and BHMF contain key information on the physics characterizing the radiative 
and accretion processes of high-z SMBHs, as well as seeding mechanisms in principle.
Two approaches have been utilized to model those distribution functions.  
The first is semi-analytical modeling, in which BH seed formation, gas accretion, and BH mergers associated 
with the hierarchical growth of the parent dark-matter halos are taken into account in a simplified way. 
This is an effective way of examining the statistical properties of BH populations and making predictions 
that can be directly compared with the QLF and BHMF
\citep[e.g.,][]{1998ApJ...503..505H,2010ApJ...718..231S,2018MNRAS.474.1995R,2018MNRAS.481.3278R,
2021MNRAS.508.2706Y,2021ApJ...910L..11K,2022MNRAS.511..616T}. 
However, due to the limitation of capturing detailed properties of the relevant physics, 
a large number of model parameters need to be calibrated based on the observed distribution functions 
of lower-$z$ quasar population \citep[e.g.,][]{2007ApJ...669...45H}
and empirical correlations between the SMBH mass and their host galaxy properties seen
in the local universe \citep{2013ARA&A..51..511K}.
The second approach is more phenomenological, using extrapolation of the low-$z$ QLF evolution with 
fitting function forms \citep[e.g.,][]{2020MNRAS.495.3252S,2022arXiv220702233F}.
This method also allows us to forecast the higher-$z$ QLF and bring insights for future surveys.
Nevertheless, those semi-analytical and phenomenological approaches introduce numerous physical parameters
and make the key mechanisms reproducing the observed results obscure.
Therefore, it is required to construct a theoretical model that is characterized by a minimum number of parameters
but capable of dealing with essential physical processes.

To determine the initial conditions of the early BH assembly,
the formation pathway of seed BHs has been extensively investigated by numerical simulations and semi-analytical studies
%Formation of seed BHs is the starting point to characterize the early assembly of massive BH populations.
%Their formation pathway and expected mass range have been extensively investigated by numerical simulations and
%semi-analytical studies
\citep{2006MNRAS.370..289B,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T}.
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
and (iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the heavy-seed formation
can be accelerated in the rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and the frequency of halo mergers.
As a result, \citet{2021ApJ...917...60L} predicted a top-heavy mass distribution of seed BHs ranging from $10^2-10^5~\Msun$.

In this paper, we propose a theoretical model for the redshift-dependent QLF and BHMF at $z\gtrsim 6$,
% applying the initial mass function (IMF) of seed BHs obtained from \cite{2021ApJ...917...60L}.
applying the semi-analytic seed formation model from \cite{2021ApJ...917...60L} to study the BH growth in common regions in the early universe.
We assume that the early BH population experiences multiple accretion bursts, in each of which the Eddington ratio is assigned 
following a Schechter distribution function. 
We conduct the Markov Chain Monte Carlo (MCMC) fitting procedure to optimize the BH growth parameters 
so that the observed BHMF and QLF at $z\simeq 6$ are simultaneously reproduced. 
Our best-fit model suggests several episodes of modest super-Eddington accretion and requires
individual burst durations of $\simeq 20-30$ Myr to avoid over(under)-production of BHs at the high-mass end.
Furthermore, we construct the redshift-dependent QLF and BHMF that imply a rapid decay of 
their number and mass toward higher redshifts of $z>6$.
We also find that the observed Eddington-ratio distribution function is skewed to a log-normal shape from the intrinsic Shechter function
owing to detection limits of quasar observations.
Those results will be unveiled by future deep and wide surveys with the James Webb Space Telescope, Roman Space Telescope, and Euclid.

\if0
To construct the picture of BH evolution, the initial mass function (IMF) of BH seeds is required.
Studies have shown that Pop III stars (or the seed BH population) formed in mini-halos is top-heavy, while capped with $\sim 10^3~\Msun$ 
\citep{2014ApJ...781...60H,2015MNRAS.448..568H}.
Focusing on the progenitor halos of supermassive BHs at $z\sim$ 6, 
our recent semi-analytical work produces an IMF of BH seeds ranging from $10^2-10^5~\Msun$ \citep{2021ApJ...917...60L}. 
The IMF is consistent with the three-dimensional radiation hydrodynamical simulations 
constituting the stellar evolution with large scale mass inflow under realistic halo circumstances (Toyouchi et al. 2022).
In this work, we apply the seeding model to examine the BH growth in common regions in the early universe, 
constrained by the observable QLF and BHMF down to $z\sim$ 6, 
% our aim is to understand the growth patterns of early BHs.
\fi
